Methods and Results

### 2022-10-12 ms.Rmd
Revised MS for Pau et al

#### Meta-analysis
We analysed effect size data using the *metafor* @Metafor (vers. `r utils::packageVersion("metafor")`) package in R @R (vers. `r paste0(R.Version()$major, ".", R.Version()$minor)`). To estimate the overall effect of male-harm on female fitness we first fit multi-level meta-analytic models (intercept only) @nakagawa2012methodological. We plotted overall meta-analytic means using the orchard plot package @NakaOrchard (vers. `r utils::packageVersion("orchard")`). Overall, we collected between `r n_eff_study[1]` to `r n_eff_study[2]` effect sizes per study. To control for sources of non-independence, we included study and species-level random effects @Noble2017. We also fit a model that included a species-level random effect with a phylogenetic correlation matrix derived using the Open Tree of Life Database @rotl (vers. `r utils::packageVersion("rotl")`). We used Grafen's method @Grafen to calculate branch lengths using the R package *ape* @ape (vers. `r utils::packageVersion("ape")`). Overall, a model containing only a species-level random effect variance was better supported than a model that estimated a phylogenetic variance ($\Delta_{AIC_{c}}$ = `r dAICc`). As such, we did not include phylogeny in our models.

@@

Using our MLMA models we also calculated effect size heterogeneity using $I^2$ [@HigginsThompson; @nakagawa2012methodological]. $I^2_{study}$ and $I^2_{species}$ are defined as the proportion of effect size variance over total variance as a result of between study and species effects, respectively.In contrast, $I^2_{total}$ was calculated as the proportion of total effect size variance after accounting for, or removing, total sampling variance.

We explored drivers of effect size heterogeneity using multi-level meta-regression (MLMR) models. Our models included fixed effects (i.e., moderators) that we *a priori* predicted would impact female fitness. More specifically, we predicted that the intensity of sexual selection and sexual conflict would affect how much harm males did to females. To test these predictions, we included an index of sexual size dimorphism (SSD), which involved taking the log transformed ratio between male body size to female body size for the species (i.e.,  $log \left( \frac{Male_{body size}}{Female_{bodysize}} \right)$, hereafter referred to as SSD). Positive SSD values indicate species with larger males compared to females, whereas negative SSD values indicate species where females are larger than males. While this does not directly measure the intensity of sexual selection it is an intuitive, interpretable metric that is normally distributed around zero [@Harrison2022; @Lovitch1992].

In addition to SSD, we also categorised effect sizes as having come from species with different types of male harm and sperm competition intensity (see moderators section). These included species with direct male harm (i.e., species with traumatic insemination) as well as species with indirect male harm (i.e., species that harm females through harassment). There are also a number of species where it is clear both direct and indirect forms of male harm exist. We also fit separate models to a two-level categorical moderator describing the sperm competition intensity of the species (i.e, "high" or "low"). In both these models, we controlled for SSD index, but included z-transformed SSD, such that meta-analytic means in each categorical moderator are for species of average SSD. We also accounted for the possibility that SSD might vary depending on the specific harm type and sperm competition intensity (i.e., interaction between SSD and Harm type and Sperm competition intensity). We evaluated support for interactions by comparing a model including the interaction to a model without it, comparing $AIC_{c}$ between the two models. We chose the model with the lowest $AIC_{c}$ if they differed by greater than 2, otherwise, we choose the most parsimonious model. All models comparing fixed effects were fit with maximum likelihood for model comparison of fixed effect structure, and then subsequently re-fit with restricted maximum likelihood when the fixed effect structure was identified.

@@

```{r funnel, fig.height=7.62500, fig.width=14.90278, fig.cap="A) Funnel plot of effect size as a function of precision (i.e., inverse of sampling standard error). B) orchaRd plot of the overall meta-analytic mean efect size, 95% confidnece intervals (thick black bars) and 95% prediction intervals (whiskers)"}
p1 <- ggplot(data, aes(x = SMDH, y = 1 / sqrt(v_SMDH), xmin = -6, xmax = 6)) +
geom_point(aes(colour = class), size = 3) +
labs(x = "SMDH", 
	 y = TeX("Precision $\\left(\\frac{1}{\\sqrt{v_{SMDH}}}\\right)$"),
	 colour = "Class") +
scale_color_grey() + 
#scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9")) + 
theme_classic()

@@

Male harm type appeared to impact the overall magnitude of effects explaining `r male_harm_r2[1]*100`% of effect size variance (Figure \@ref(fig:maleharm)). Species where mating results in both direct and indirect harm to females have males that significantly impact female fitness (overall meta-analytic mean: `r robust_model1.5_2$beta[1]`, 95% CI: `r robust_model1.5_2$ci.lb[1]` to `r robust_model1.5_2$ci.ub[1]`,  p = `r p_value(robust_model1.5_2$pval[1])`). Surprisingly, however, species with direct female harm only (i.e., traumatic insemination), showed a small and opposite effect on female fitness (overall meta-analytic mean: `r robust_model1.5_2$beta[2]`, 95% CI: `r robust_model1.5_2$ci.lb[2]` to `r robust_model1.5_2$ci.ub[2]`; Figure \@ref(fig:maleharm)), however, this effect did not differ significantly from zero (p = `r p_value(robust_model1.5_2$pval[2])`). While indirect male harm also resulted in a smaller overall meta-analytic mean relative to species with both, it was still significantly positive (overall meta-analytic mean: `r robust_model1.5_2$beta[3]`, 95% CI: `r robust_model1.5_2$ci.lb[3]` to `r robust_model1.5_2$ci.ub[3]`,  p = `r p_value(robust_model1.5_2$pval[3])`), and only marginally different than species with both direct and indirect male harm (Figure \@ref(fig:maleharm)). Regardless, differences among species exhibiting different types of female harm seemed to be driven primarily by a single study (Taylor, 2008) on *Drosophila melanogaster* (very high Cook's distance). Removing this effect resulted in no significant difference between harm type categories.

Harm type did not impact the overall variance in female fitness between control and treatment groups ($F_{2,140}$ = `r model1_lnCVR_harm$QM`, p = `r model1_lnCVR_harm$QMp` - Figure \@ref(fig:maleharm)).

@@

The lnCVR effect size does have limitations which are not always upheld in practice [@nakagawa2015meta]. As such, we tested whether these assumptions impacted our inferences by re-running analyses using the log Standard Deviation (lnSD) as our effect size while controlling for log Mean (lnX) using a multilevel meta-regression model using *metafor* (version `r utils::packageVersion("metafor")`) that included lnX along with a two level categorical predictor for treatment (Control vs Treatment) as fixed effects to both control for mean-variance relationship and estimate the effect contrast between control and treatment groups. The effect contrast should approximate lnCVR if assumptions are upheld. As random effects, we included study, species, and a residual variance (given *metafor* does not include this by default). We also included a random slope for each control and treatment comparison within a study to model variation in effects across studies.

>>>>

#### Meta-analysis
We analysed effect size data using the *metafor* @Metafor (vers. `r utils::packageVersion("metafor")`) package in R @R (vers. `r paste0(R.Version()$major, ".", R.Version()$minor)`). To estimate the overall effect of male-harm on female fitness (for both mean and variance) we first fit multi-level meta-analytic models (intercept only) @nakagawa2012methodological. We plotted overall meta-analytic means using the orchard plot package @NakaOrchard (vers. `r utils::packageVersion("orchard")`). Overall, we collected between `r n_eff_study[1]` to `r n_eff_study[2]` effect sizes per study. To control for sources of non-independence, we included study and species-level random effects @Noble2017. We also fit a model that included a species-level random effect with a phylogenetic correlation matrix derived using the Open Tree of Life Database @rotl (vers. `r utils::packageVersion("rotl")`). We used Grafen's method @Grafen to calculate branch lengths using the R package *ape* @ape (vers. `r utils::packageVersion("ape")`). Overall, a model containing only a species-level random effect variance was better supported than a model that estimated a phylogenetic variance ($\Delta_{AIC_{c}}$ = `r dAICc`). As such, we did not include phylogeny in our models.

@@

Using our MLMA models we also calculated effect size heterogeneity using $I^2$ [@HigginsThompson; @nakagawa2012methodological]. $I^2_{study}$ and $I^2_{species}$ are defined as the proportion of effect size variance over total variance as a result of between study and species effects, respectively. In contrast, $I^2_{total}$ was calculated as the proportion of total effect size variance after accounting for, or removing, total sampling variance.

We explored drivers of effect size heterogeneity using multi-level meta-regression (MLMR) models. Our models included fixed effects (i.e., moderators) that we *a priori* predicted would impact female fitness or variance in female fitness. More specifically, we predicted that the intensity of sexual selection and sexual conflict would affect how much harm males did to females (mean fitness only). To test these predictions, we included an index of sexual size dimorphism (SSD), which involved taking the log transformed ratio between male body size to female body size for the species (i.e.,  $log \left( \frac{Male_{body size}}{Female_{bodysize}} \right)$, hereafter referred to as SSD). Positive SSD values indicate species with larger males compared to females, whereas negative SSD values indicate species where females are larger than males. While this does not directly measure the intensity of sexual selection it is an intuitive, interpretable metric that is normally distributed around zero [@Harrison2022; @Lovitch1992].

In addition to SSD, we also categorised effect sizes as having come from species with different types of male harm and sperm competition intensity (see moderators section). These included species with direct male harm (i.e., species with traumatic insemination) as well as species with indirect male harm (i.e., species that harm females through harassment). There are also a number of species where it is clear both direct and indirect forms of male harm exist. We also fit separate models to a two-level categorical moderator describing the sperm competition intensity of the species (i.e, "high" or "low"). In both these models, we controlled for SSD index, but included z-transformed SSD, such that meta-analytic means in each categorical moderator are for species of average SSD. We also accounted for the possibility that SSD might vary depending on the specific harm type and sperm competition intensity (i.e., interaction between SSD and Harm type and Sperm competition intensity). We fit models with all of the above moderators for SMDH, but only included the harm type moderator in lnCVR models because we had no clear *a priori* predictions about how variance in female fitness should be affected by SSD and sperm competition intensity. We evaluated support for interactions by comparing a model including the interaction to a model without it, comparing $AIC_{c}$ between the two models. We chose the model with the lowest $AIC_{c}$ if they differed by greater than 2, otherwise, we choose the most parsimonious model. All models comparing fixed effects were fit with maximum likelihood for model comparison of fixed effect structure, and then subsequently re-fit with restricted maximum likelihood when the fixed effect structure was identified.

@@

```{r funnel, fig.height=7.62500, fig.width=14.90278, fig.cap="A) Funnel plot of effect size as a function of precision (i.e., inverse of sampling standard error). B) orchaRd plot of the overall meta-analytic mean SMDH. C) orchard plot of overall meta-analytic mean for lnCVR. In B & C the mean is presented with 95% confidence intervals (thick black bars) and 95% prediction intervals (whiskers)."}
p1 <- ggplot(data, aes(x = SMDH, y = 1 / sqrt(v_SMDH), xmin = -6, xmax = 6)) +
geom_point(aes(colour = class), size = 3) +
labs(x = "SMDH", 
	 y = TeX("Precision $\\left(\\frac{1}{\\sqrt{v_{SMDH}}}\\right)$"),
	 colour = "Class") +
scale_color_grey() + 
#scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9")) + 
theme_classic()

@@

Male harm type appeared to impact the overall magnitude of effects explaining `r male_harm_r2[1]*100`% of effect size variance (Figure \@ref(fig:maleharm)). Species where mating results in both direct and indirect harm to females have males that significantly impact female fitness (overall meta-analytic mean: `r robust_model1.5_2$beta[1]`, 95% CI: `r robust_model1.5_2$ci.lb[1]` to `r robust_model1.5_2$ci.ub[1]`,  p = `r p_value(robust_model1.5_2$pval[1])`). Surprisingly, however, species with direct female harm only (i.e., traumatic insemination), showed a small and opposite effect on female fitness (overall meta-analytic mean: `r robust_model1.5_2$beta[2]`, 95% CI: `r robust_model1.5_2$ci.lb[2]` to `r robust_model1.5_2$ci.ub[2]`; Figure \@ref(fig:maleharm)), however, this effect did not differ significantly from zero (p = `r p_value(robust_model1.5_2$pval[2])`). While indirect male harm also resulted in a smaller overall meta-analytic mean relative to species with both, it was still significantly positive (overall meta-analytic mean: `r robust_model1.5_2$beta[3]`, 95% CI: `r robust_model1.5_2$ci.lb[3]` to `r robust_model1.5_2$ci.ub[3]`,  p = `r p_value(robust_model1.5_2$pval[3])`), and only marginally different than species with both direct and indirect male harm (Figure \@ref(fig:maleharm)). Regardless, differences among species exhibiting different types of female harm seemed to be driven primarily by a single study (Taylor, 2008) on *Drosophila melanogaster* (very high Cook's distance). Removing this effect resulted in no significant difference between harm type categories. Harm type did not impact the overall variance in female fitness between control and treatment groups ($F_{2,140}$ = `r model1_lnCVR_harm$QM`, p = `r model1_lnCVR_harm$QMp` - Figure \@ref(fig:maleharm)).

@@

The lnCVR effect size does have limitations which are not always upheld in practice [@nakagawa2015meta]. As such, we tested whether these assumptions impacted our inferences by re-running analyses using the log Standard Deviation (lnSD) as our effect size while controlling for log Mean (lnX) using a multilevel meta-regression model using *metafor* (version `r utils::packageVersion("metafor")`) that included lnX along with a two-level categorical predictor for treatment (Control vs Treatment) as fixed effects to both control for mean-variance relationship and estimate the effect contrast between control and treatment groups. The effect contrast should approximate lnCVR if assumptions are upheld. As random effects, we included study, species, and a residual variance (given *metafor* does not include this by default). We also included a random slope for each control and treatment comparison within a study to model variation in effects across studies.
